---
title: "Prioritizing potential aquaculture"
author: "Marina Kochuten"
date: last-modified
execute: 
  warning: false
  message: false
format: html
editor_options: 
  chunk_output_type: console
---

## Project description

For this project, I am determining which Exclusive Economic Zones (EEZ) on the West Coast of the US are best suited to developing marine aquaculture for several species of oysters and dungeness crab. Marine aquaculture has the potential to play an important role in the global food supply as a more sustainable protein option than land-based meat production (Hall et al. 2011). Gentry et al. mapped the potential for marine aquaculture globally based on multiple constraints, including ship traffic, dissolved oxygen, and bottom depth. They found that global seafood demand could be met using less than 0.015% of the global ocean area (Gentry et al. 2017). Suitable locations are determined based on range of suitable sea surface temperature (SST) and depth values for the species.

## Data details

#### Suitable Growing Conditions

**Oysters**
- sea surface temperature: 11-30°C
- depth: 0-70 meters below sea level

**Dungeness Crab**
- sea surface temperature: 3-19°C
- depth: 0-360 meters below sea level

#### Sea Surface Temperature

I use average annual sea surface temperature (SST) from the years 2008 to 2012 to characterize the average sea surface temperature within the region. The data we are working with was originally generated from NOAA’s 5km Daily Global Satellite Sea Surface Temperature Anomaly v3.1.

#### Bathymetry

To characterize the depth of the ocean, I use the General Bathymetric Chart of the Oceans (GEBCO).

#### Exclusive Economic Zones

I designate maritime boundaries using Exclusive Economic Zones off of the west coast of US from Marineregions.org

## Setup

```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Load libraries ----
library(sf)
library(terra)
library(tidyverse)
library(tmap)
library(here)
library(testthat)
library(kableExtra)
```

Load the data:
```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Read in data ----

# Sea surface temp
sst_2008 <- rast(here("data", "average_annual_sst_2008.tif"))
sst_2009 <- rast(here("data", "average_annual_sst_2009.tif"))
sst_2010 <- rast(here("data", "average_annual_sst_2010.tif"))
sst_2011 <- rast(here("data", "average_annual_sst_2011.tif"))
sst_2012 <- rast(here("data", "average_annual_sst_2012.tif"))

# Bathymetry
depth <- rast(here("data", "depth.tif"))

# Exclusive economic zones for maritime boundaries
wc_regions <- read_sf(here("data", "wc_regions_clean.shp"))
```

Prepare data for processing:
```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Combine SST rasters into a stack ----
sst_all <- c(sst_2008,
             sst_2009,
             sst_2010,
             sst_2011,
             sst_2012)

# Match CRSs ----

# Regions
wc_regions <- st_transform(wc_regions, crs = "EPSG:4326")

# Depth
depth <- depth |>
  project("EPSG:4326")

# SST
sst_all <- sst_all|>
  project("EPSG:4326")

# Check that CRS match
test_that("All CRSs match", 
          {expect_true(st_crs(wc_regions) == st_crs(depth) & st_crs(wc_regions) == st_crs(sst_all))
})
```

Process data
```{r}
#| code-fold: true
#| code-summary: "Show the code"

# Create raster with mean SST from 2008-2012 ----
sst_mean <- mean(sst_all, na.rm = TRUE)  # Remove NA values

# Convert mean SST from Kelvin to Celsius ----
sst_mean <- sst_mean - 273.15

# Crop depth to match the extent of mean SST ----
ext(depth) <- ext(sst_mean)

# Resample the depth data to match the resolution of the SST data using the nearest neighbor approach ----
depth <- resample(depth, sst_mean, method = "near")

# Stack depth and mean SST to check that resolution, extent, and CRS match
sst_depth <- c(sst_mean,
               depth)

```

Find suitable locations for marine aquaculture
```{r}
#| code-fold: true
#| code-summary: "Show the code"
#| output: false
#| collapse: true

# Reclassify SST and depth into locations that are suitable for oysters ----

# Create reclassification matrix - sst oysters
rcl_sst_oysters <- matrix(c(-Inf, 11, 0,    # unsuitable
                            11, 30, 1,      # suitable
                            30, Inf, 0),    # unsuitable
                            ncol = 3, byrow = TRUE)

# Reclassify SST
sst_oysters <- classify(sst_mean, rcl = rcl_sst_oysters)
plot(sst_oysters)

# Create reclassification matrix - depth oysters
rcl_depth_oysters <- matrix(c(-Inf, -70, 0,   # unsuitable
                              -70, 0, 1,      # suitable
                              0, Inf, 0),     # unsuitable
                              ncol = 3, byrow = TRUE)

# Reclassify depth
depth_oysters <- classify(depth, rcl = rcl_depth_oysters)
plot(depth_oysters)

# Find locations that satisfy both SST and depth conditions for oysters ----
oysters_suitable <- lapp(c(sst_oysters, depth_oysters), fun = function(sst, depth) sst * depth) # multiplication will only place 1 in areas that have 1 for both sst and depth

# Set unsuitable areas as NA
oysters_suitable[oysters_suitable == 0] <- NA
plot(oysters_suitable)
```

Determine the most suitable EEZ
```{r}
# Mask suitable oysters with wc_regions to crop suitable cells to within EEZs ----
oysters_masked <- mask(oysters_suitable, wc_regions)
plot(oysters_masked)

# Rasterize wc_regions for zonal operations ----
eez_rast <- rasterize(wc_regions, oysters_masked, field = "rgn_id")
plot(eez_rast)

# Find area of each grid cell ----
cell_size <- cellSize(oysters_masked, unit = "km") # This is not giving an output I would have expected
plot(cell_size)

# Find total number of cells within each EEZ ----
eez_oyster_area <- zonal(cell_size * oysters_masked, eez_rast, fun = "sum", na.rm = TRUE) |>
  arrange(area)

# Make it look preety with spData us states
```





## Citations

#### Data

**Suitable Growing Conditions**
Palomares, M.L.D. and D. Pauly. Editors. 2024. SeaLifeBase. World Wide Web electronic publication. www.sealifebase.org, version (08/2024). Accessed 2024-11-09

**Sea Surface Temperature**
NOAA Coral Reef Watch. 2018, updated daily. NOAA Coral Reef Watch Version 3.1 5km Daily Global Satellite Sea Surface Temperature Anomaly Product, 2008-2012. College Park, Maryland, USA: NOAA Coral Reef Watch. Data set accessed 2024-11-09 at https://coralreefwatch.noaa.gov/product/5km/index_5km_ssta.php.

**Bathymetry**
GEBCO Compilation Group (2024) GEBCO 2024 Grid (doi:10.5285/1c44ce99-0a0d-5f4f-e063-7086abc0ea0f). Data set accessed 2024-11-09 at https://www.gebco.net/data_and_products/gridded_bathymetry_data/#area

**Exclusive Economic Zones**
Marine Regions (2024) Flanders Marine Institute. Data set accessed 2024-11-09 at https://www.marineregions.org/eez.php. 

#### Literature

Hall, S. J., Delaporte, A., Phillips, M. J., Beveridge, M. & O’Keefe, M. Blue Frontiers: Managing the Environmental Costs of Aquaculture (The WorldFish Center, Penang, Malaysia, 2011).

Gentry, R. R., Froehlich, H. E., Grimm, D., Kareiva, P., Parke, M., Rust, M., Gaines, S. D., & Halpern, B. S. Mapping the global potential for marine aquaculture. Nature Ecology & Evolution, 1, 1317-1324 (2017).